---
title: "Heatwaves"
author: "Gregor Steiner"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
output: pdf_document
bibliography: references.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


Using data provided by the US Climate Reference Network [@Diamond_2013], I identify heatwaves. There are different ways to define heatwaves. Following the National Climate Data Center (NCDC), I use the following definition: A heatwave consists of two or more consecutive days in which the average, minimum or maximum temperature exceeds the 85th percentile of July and August temperatures [@Habeeb_2015]. 

```{r}


########## Daily Data ##########
dat <- do.call(rbind, Map(function(y){
  # create url
  url <- paste0("https://www.ncei.noaa.gov/pub/data/uscrn/products/daily01/", y, "/CRND0103-", y, "-AZ_Tucson_11_W.txt")
  
  # read data (documentation: https://www.ncei.noaa.gov/pub/data/uscrn/products/daily01/)
  dat <- read.table(url)
  
  # create dataframe
  dat.out <- data.frame(Date = as.Date(as.character(dat[, 2]), "%Y%m%d"),
                        AvgTemp = ifelse(dat[, 9] != -9999, dat[, 9], NA),
                        Name = "Tucson",
                        Longitude = dat[, 4],
                        Latitude = dat[, 5])
  
  # drop temperature NAs
  dat.out <- dat.out[!is.na(dat.out[, "AvgTemp"]), ]
  
  # return
  return(dat.out)
  
}, 2002:2021))


# Indicate Heatwaves
# compute threshold (85th percentile of july and august temperatures)
bool.jul.aug <- format(dat$Date, "%m") %in% c("07", "08") 
threshhold <- quantile(dat[bool.jul.aug, "AvgTemp"], 0.85, na.rm = TRUE)

# Mark as heatwave all consecutive days that exceed the threshhold
bool <- dat[, "AvgTemp"] > threshhold
dat$Heatwave <- 0

for (i in 1:(nrow(dat)-1)) {
  if(all(bool[i:(i+1)])){
    dat[i:(i+1), "Heatwave"] <- 1
  }
}



```



# Visualization

```{r}
# Get dates with exte´reme heat
heatdates <- dat[dat$Heatwave == 1, "Date"]

# make graph
plot(dat$Date, dat$AvgTemp, type = "n",
     ylab = "Average Temperature", xlab = "")

abline(v = heatdates, col = rgb(1, 0, 0, 0.3), lwd = 3)
lines(dat$Date, dat$AvgTemp, lwd = 1)





```



# References



