---
title: "Heatwaves"
author: "Gregor Steiner"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
output: pdf_document
bibliography: references.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.pos = "center",
                      fig.width = 8)
```


# Daily Data

Using data provided by the US Climate Reference Network [@Diamond_2013], I identify extreme heat events in the historical temperature time series. Following the National Climate Data Center (NCDC), I define an extreme heat event as two or more consecutive days in which the average temperature exceeds the 85th percentile of July and August temperatures [@Habeeb_2015]. 

```{r}


########## Daily Temperature Data ##########
dat.stations <- openxlsx::read.xlsx("weather_stations.xlsx")

# map over stations and row bind everything together
dat <- do.call(rbind, Map(function(station, name){
  
  # map over years and row bind
  dat <- do.call(rbind, Map(function(y){
    # create url
    url <- paste0("https://www.ncei.noaa.gov/pub/data/uscrn/products/daily01/", y, "/CRND0103-", y, "-", station, ".txt")
    
    # read data (documentation: https://www.ncei.noaa.gov/pub/data/uscrn/products/daily01/)
    dat <- read.table(url)
    
    # create dataframe
    dat.out <- data.frame(Date = as.Date(as.character(dat[, 2]), "%Y%m%d"),
                          AvgTemp = ifelse(dat[, 9] != -9999, dat[, 9], NA),
                          Name = name,
                          Longitude = dat[, 4],
                          Latitude = dat[, 5])
    
    # drop temperature NAs
    dat.out <- dat.out[!is.na(dat.out[, "AvgTemp"]), ]
    
    # return
    return(dat.out)
    
  }, 2003:2021))
  
  
  # Indicate Heatwaves
  # compute threshold (85th percentile of july and august temperatures)
  bool.jul.aug <- format(dat$Date, "%m") %in% c("07", "08") 
  threshhold <- quantile(dat[bool.jul.aug, "AvgTemp"], 0.85, na.rm = TRUE)
  
  # Mark as heatwave all consecutive days that exceed the threshhold
  bool <- dat[, "AvgTemp"] > threshhold
  dat$Heatwave <- 0
  
  for (i in 1:(nrow(dat)-1)) {
    if(all(bool[i:(i+1)])){
      dat[i:(i+1), "Heatwave"] <- 1
    }
  }
  
  return(dat)

},
dat.stations[, "Station"],
paste(dat.stations[, "City"], dat.stations[, "State"], sep = ", ")))

```




```{r fig.cap = "Daily average temperatures with extreme heat events shaded in red"}
par(mfrow = c(round(nrow(dat.stations) / 2), 2),
    mar = c(2, 4, 2, 1))

invisible(lapply(unique(dat$Name), function(name){
  
  dat.int <- dat[dat$Name == name, ]  
  
  # Get dates with extreme heat
  heatdates <- dat.int[dat.int$Heatwave == 1, "Date"]
  
  # make graph
  plot(dat.int$Date, dat.int$AvgTemp, type = "n",
       ylab = "Temperature in °C", xlab = "", main = name)
  
  abline(v = heatdates, col = rgb(1, 0, 0, 0.3), lwd = 1)
  lines(dat.int$Date, dat.int$AvgTemp, lwd = 1)

}))


```






```{r}
########## Fatal Crash Data ##########
dat.crash <- haven::read_dta("20140100_indiv_records.dta")

# only keep relevant dates
dat.crash <- dat.crash[dat.crash$date >= as.Date("2003-01-01"), ]

# aggregate by county and state
crash.agg <- aggregate(list(Cases = dat.crash$st_case),
                       by = list(
                         Date = dat.crash$date,
                         State = dat.crash$state,
                         County = dat.crash$county
                         ),
                       FUN = length)
crash.agg <- crash.agg[order(crash.agg$State, crash.agg$County, crash.agg$Date), ]


```



# References



