---
title: "Extreme Heat Events and Fatal Car Accidents"
author: "Gregor Steiner"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: "TeX files/references.bib" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.pos = "center",
                      fig.width = 8,
                      fig.height = 4)
```




```{r}
# preliminaries
library(data.table)
library(usmap)
library(ggplot2)

dat <- readRDS("Code & Data/Data.RDS")

```


# Natural Disasters and Education

In my thesis I explore the link between natural disasters and education outcomes. The Stanford Education Data Archive [@SEDA] provides summary statistics for standardized test scores by county, year, grade and subject. Data on natural disasters is available at the FEMA database and the **rfema** package [@rfema] provides easy access.

Below you can see the number of cumulative disasters from 2009 to 2018 by county.

```{r}
plot_usmap(data = dat[year == max(year) & grade == 3 & subject == "mth"], values = "CumuDisasters") +
  scale_fill_viridis_c(name = "Disasters")


```








## References


