---
title: "Natural Disasters and Education"
author: "Gregor Steiner"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: "TeX files/references.bib" 
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.pos = "center",
                      fig.width = 8,
                      fig.height = 4)
```




```{r}
# preliminaries
library(data.table)
library(usmap)
library(ggplot2)

dat <- readRDS("Code & Data/Data.RDS")

```


# Natural Disasters and Education

In my thesis I explore the link between natural disasters and education outcomes. The Stanford Education Data Archive [@SEDA] provides standardized test scores by county, year, grade and subject. Data on natural disasters is available at the FEMA database and the **rfema** package [@rfema] provides easy access.

The outcomes of interest are mean test scores by county, and mean test scores for black, hispanic, female, and economically disadvantaged students.

```{r}
dat.summary <- dat[, .("Mean test score" = cs_mn_all,
                       "Mean test score (black students)" = cs_mn_blk,
                       "Mean test score (hispanic students)" = cs_mn_hsp,
                       "Mean test score (female students)" = cs_mn_fem,
                       "Mean test score (economically disadvantaged students)" = cs_mn_ecd)]

vtable::sumtable(dat.summary)
```

To identify a causal effect, I use an event study design. Due to likely very heterogenous treatment effects, I employ the estimator by @Sun_2021.



## References


