---
title: "Extreme Heat Events and Fatal Car Accidents"
author: "Gregor Steiner"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: "TeX files/references.bib" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.pos = "center",
                      fig.width = 8,
                      fig.height = 6)
```




```{r}
# preliminaries
library(data.table)

```

## Extreme Heat Events and Fatal Car Accidents

In my master thesis I want to explore the causal effect of extreme heat events on fatal car accidents.

### Weather Data

Using data provided by the Daily Global Historical Climatology Network [@Menne_2012], I identify extreme heat events in the historical temperature time series. Historical weather data is available for about 1200 measurement stations in the US. Based on the latitude and longitude coordinates I assign each station to the county it belongs to. In case there is more than one station per county, I take the mean among those.

Following the National Climate Data Center (NCDC), I define an extreme heat event as days in which the maximum or minimum temperature exceed the 85th percentile of July and August maximum or minimum temperatures [@Habeeb_2015]. Below you can see the average (across all counties) number of heat days based on minimum and maximum temperatures for each year.


```{r}
# read data
data <- readRDS("Code & Data/DataCombined.RDS")


# # Define counties of interest
# counties <- c("JACKSON COUNTY, OREGON", "ONTARIO COUNTY, NEW YORK",
#              "ORANGE COUNTY, CALIFORNIA", "YORK COUNTY, SOUTH CAROLINA",
#              "PIMA COUNTY, ARIZONA", "CHAMPAIGN COUNTY, ILLINOIS")
# 
# 
# # Make plots
# par(mfrow = c(round(length(counties) / 2), 2),
#     mar = c(2, 4, 2, 1))
# invisible(lapply(counties, function(x){
#   # Select County
#   dat <- data[data$StateCounty == x, ]
#   
#   # Get dates with extreme heat
#   heatdates <- unlist(dat[dat$EHE.MAX == 1, "DATE"])
#   
#   # make graph
#   plot(dat$DATE, dat$TMAX, type = "n",
#        ylab = "Temperature in °C", xlab = "", main = x)
#   
#   abline(v = heatdates, col = rgb(1, 0, 0, 0.3), lwd = 1)
#   lines(dat$DATE, dat$TMAX, lwd = 1)
#   
# }))

```


```{r fig.height=3}
# number of extreme heat events by year
setDT(data)

# display average number of ehe days in a year, based on both minimum and maximum temperature
EHEs.yearly <- data[
  , .(SUM.EHE.MAX = sum(EHE.MAX), SUM.EHE.MIN = sum(EHE.MIN)), by = .(YEAR = format(DATE, "%Y"), StateCounty)
  ][
    , .(AVG.EHE.MAX = mean(SUM.EHE.MAX), AVG.EHE.MIN = mean(SUM.EHE.MIN)), by = .(YEAR)
    ]

# plot results
par(mar = c(2.5, 4, 1, 1))
invisible(within(EHEs.yearly, {
  plot(YEAR, AVG.EHE.MAX, type = "n",
       xlab = "", ylab = "Number of extreme heat days")
  grid()
  lines(YEAR, AVG.EHE.MAX, col = 2, lwd = 2)
  lines(YEAR, AVG.EHE.MIN, col = 4, lwd = 2)
  legend("topleft", fill = c(2, 4), legend = c("Maximum", "Minimum"))
}))

```


### Car Crash Data

Data on fatal car accidents is available from the [Fatality Analysis Reporting System (FARS)](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars). Below you can see the number of fatal car accidents for each year (on a national level).

```{r fig.height=3}
# plot number of fatal car accidents
FatAccs <- data[, .(FatAccs = sum(FatalAccidents)), by = .(YEAR = format(DATE, "%Y"))]

par(mar = c(2.5, 4, 1, 1))
invisible(within(FatAccs, {
  plot(YEAR, FatAccs, type = "n",
       xlab = "", ylab = "Fatal car accidents")
  grid()
  lines(YEAR, FatAccs, col = 4, lwd = 2)
}))

```


### Causal Effect?

A naive point estimate of the average treatment effect using the difference of conditional means (conditional on treatment) indicates positive causal effects, that is extreme heat increases the number of fatal car accidents.

```{r}
# reconvert to data.frame to get rid of DT functionality
data <- as.data.frame(data)

knitr::kable(sapply(c("EHE.MAX", "EHE.MIN"), function(x){
  # condition on treatment
  treated <- data[, x] == 1
  
  # estimate tau
  tau.hat <- mean(data[treated, "FatalAccidents"]) - mean(data[!treated, "FatalAccidents"])
  
  return(tau.hat)
}) |> setNames(c("MaximumTemperature", "MinimumTemperature")), col.names = "ATE")


```
However, this estimator is only useful if there is no selection bias and if treatment effects are homogenous, which is unlikely. Therefore, my next step will be working on a credible identification strategy.





## References



