---
title: "Natural Disasters and Education"
author: "Gregor Steiner"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
output:
  md_document:
    variant: markdown_github
bibliography: "TeX files/references.bib" 
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.pos = "center",
                      fig.width = 8,
                      fig.height = 4)
```




```{r}
# preliminaries
library(data.table)
library(usmap)
library(ggplot2)
library(fixest)

source("Code & Data/00_AuxFunctions.R")

dat <- setDT(readRDS("Code & Data/Data.RDS"))

```


# Natural Disasters and Education

In my thesis I explore the link between natural disasters and education outcomes. The Stanford Education Data Archive [@SEDA] provides standardized test scores by county, year, grade and subject. Data on natural disasters is available at the FEMA database and the **rfema** package [@rfema] provides easy access. The map below shows natural disaster exposure by county for schoolyears 2008-09 through 2017-2018.

```{r}

# plot cumulative disasters
dat[, CumuDisasters := cumsum(Disasters), by = .(fips, grade, subject)]

fema.cum <- dat[, .(Disasters = CumuDisasters[!is.na(CumuDisasters)][.N]), by = fips]
fema.cum[, DisastersGrouped := factor(cut(Disasters, breaks = c(-Inf, 0, 2, 5, 10, Inf)), labels = c("0", "1-2", "3-5", "6-10", ">10"))]

plot_usmap(data = fema.cum, values = "DisastersGrouped") +
  scale_fill_manual(values = viridisLite::viridis(length(levels(fema.cum$DisastersGrouped))),
                    name = "") +
  theme(legend.position = "right",
        legend.key.size = grid::unit(1, "cm"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        plot.margin= grid::unit(c(0,0,0,0), "mm"))

```


The outcomes of interest are mean test scores by county, and mean test scores for black, hispanic, female, and economically disadvantaged students.

```{r}

# color scheme
cols <- c("firebrick", "cornflowerblue")

# make plot
layout(matrix(c(1, 2, 3), ncol = 1, byrow = TRUE), heights = c(4, 4, 1))
par(mar = c(2, 3, 1, 1))
boxplot(dat[subject == "mth", .("Overall" = cs_mn_all,
                                "White" = cs_mn_wht,
                                "Black" = cs_mn_blk,
                                "Hispanic" = cs_mn_hsp,
                                "Female" = cs_mn_fem,
                                "Econ. disadv." = cs_mn_ecd)],
        col = cols[1], cex.axis = 1)
par(mar = c(2, 3, 1, 1))
boxplot(dat[subject == "rla", .("Overall" = cs_mn_all,
                                "White" = cs_mn_wht,
                                "Black" = cs_mn_blk,
                                "Hispanic" = cs_mn_hsp,
                                "Female" = cs_mn_fem,
                                "Econ. disadv." = cs_mn_ecd)],
        col = cols[2], cex.axis = 1)


par(mai=c(0,0,0,0))
plot.new()
legend(x = "center", legend = c("Mathematics", "RLA"),
       fill = cols, cex = 1, inset = 0, horiz = TRUE)
```

To identify a causal effect, I use an event study design. Due to likely very heterogenous treatment effects, I employ the estimator by @Sun_2021. Below you can see the results of my main analysis, indicating that natural disasters do have a negative short-term effect on academic achievement in mathematics.

```{r fig.height=8}

# fema models
model.math <- feols(c(cs_mn_all, cs_mn_wht, cs_mn_blk, cs_mn_hsp, cs_mn_fem, cs_mn_ecd) ~ 
                 sunab(TreatStart, year, ref.p = c(-1, -3000), bin.rel = c(-5:-3000)) | year + fips + grade,
               data = dat[subject == "mth"], cluster = "TreatStart")

model.rla <- feols(c(cs_mn_all, cs_mn_wht, cs_mn_blk, cs_mn_hsp, cs_mn_fem, cs_mn_ecd) ~ 
                     sunab(TreatStart, year, ref.p = c(-1, -3000), bin.rel = c(-5:-9)) | year + fips + grade,
                   data = dat[subject == "rla"], cluster = "TreatStart")



# dependent variables
dep.vars <- c("Overall", "White", "Black", "Hispanic", "Female", "Econ. Disadv.")

# plot results
layout(matrix(c(1:6, 7, 7), ncol = 2, byrow = TRUE), heights = c(4, 4, 4, 1))

par(mar = c(2, 4, 2, 1))

invis.Map(function(math, rla, name){
  
  iplot(list(math, rla), main = name, xlab = "Years to treatment",
        col = cols, ci.col = cols, pt.pch = 19,
        ci.lwd = 1, ci.width = 0.2)
  
}, model.math, model.rla, dep.vars)

par(mai=c(0,0,0,0))
plot.new()
legend(x = "center", legend = c("Math", "RLA"),
       col = cols, lwd = 2, cex = 1, inset = 0, horiz = TRUE)



```




## References


